D.I.A.G.R.A.M.: Development of Image Analysis for Graph Recognition And Modeling

### 2025-05-05 doc/article/article.tex
Draft

\section{System Architecture}
Figure~\ref{fig:pipeline} illustrates the full pipeline, composed of four main modules: the Classifier, Extractor, Transducer, and Compiler. Each module is designed to process and transform the handwritten diagram image progressively toward a structured output.

\begin{figure}[H]
\centering
\includegraphics[width=\linewidth]{example-image}
\caption{Overview of the D.I.A.G.R.A.M. system architecture.}
\label{fig:pipeline}
\end{figure}

@@

% Classifier
\subsection{Classifier Module}
TODO
\subsection{Extractor Module}
The Extractor is the key component of the D.I.A.G.R.A.M. system, able to transform the raw diagram image into a structured, category-agnostic representation.
The Extractor recognizes the diagram's components, such as shapes, lines, and text, through an object detection network, and organizes them into a unified format. This representation serves as the foundation for the subsequent Transducer module, which converts the structured data into a domain-specific markup language.
\subsection{Transducer Module}
The Transducer is responsible for converting the unified, agnostic representation of a diagram into a domain-specific markup language. 
This transformation enables the subsequent Compiler module to generate high-quality visual outputs.
\subsection{Compiler Module}
The Compiler module is the final stage of the D.I.A.G.R.A.M. pipeline. 
Its primary role is to take the structured representation of the diagram, expressed in a markup language (e.g., Mermaid.js), and generate a high-quality, editable diagram in a visual format such as PNG.

>>>>

% System Architecture
\section{System Architecture}
Figure~\ref{fig:pipeline} illustrates the full pipeline, composed of four main modules: the Classifier, Extractor, Transducer, and Compiler. Each module is designed to process and transform the handwritten diagram image progressively toward a structured output.

\begin{figure}[H]
\centering
\includegraphics[width=\linewidth]{overview.png}
\caption{Overview of the D.I.A.G.R.A.M. system architecture.}
\label{fig:pipeline}
\end{figure}

@@

\subsection{Classifier Module}
The Classifier is the first component of the pipeline. It is able to tell to which category a given image of an handwritten diagram belongs to. This is necessary in order to later know the Extractors that can process the input image.
\subsection{Extractor Module}
The Extractor is the key component of the D.I.A.G.R.A.M. system, able to transform the raw diagram image into a structured, category-agnostic representation.
The Extractor recognizes the diagram's components, such as shapes, lines, and text, through an object detection network, and organizes them into a unified format. This representation serves as the foundation for the subsequent Transducer module, which converts the structured data into a domain-specific markup language.
\subsection{Transducer Module}
The Transducer is responsible for converting the unified, agnostic representation of a diagram into a domain-specific markup language. 
This transformation enables the subsequent Compiler module to generate high-quality visual outputs.
\subsection{Compiler Module}
The Compiler module is the final stage of the D.I.A.G.R.A.M. pipeline. 
Its primary role is to take the structured representation of the diagram, expressed in a markup language (e.g., Mermaid.js), and generate a high-quality diagram in a visual format such as PNG.
